---
title: "Put gtsummary in Excel"
---

```{r set up, collapse=T}
library(here)
library(tidyverse)
library(labelled)
library(gtsummary)
library(huxtable)
library(openxlsx)
library(haven)

# Load BSA teaching data
df <- readRDS(here::here("data", "synthetic_bsa2019.rds"))

ddir <- paste0(getwd(), "/")

look_for(df)
df %>% count(skipmeal)
df %>% count(RSex)

df <- df %>% mutate(across(c(skipmeal, RSex, HEdQual3), ~fct_drop(to_factor(.x, user_na_to_na = T))))
df %>% count(skipmeal)
df %>% count(RSex)
df %>% count(HEdQual3)


# List all .R files in ddir 
file_paths <- list.files(path = ddir, pattern = "\\.R$", full.names = TRUE, all.files = T) %>% 
  map_chr(., ~str_remove(.x, ddir)) %>% 
  grep("_interim", ., value =T, invert = T) # Drop interim functions, not used.

# Set of functions to be loaded. 
file_paths

```

<br>

### Load functions
```{r, results='hide'}

# Load all functions (make sure no other .R files in this location)
map(file_paths, ~source(paste0(ddir,.x)))

```

<br> 

```{r function}

tables

```

<br>

Typical theme for gtsummary()

```{r set theme}

# For all I use the compact theme.
theme_gtsummary_compact()

```

<br> 

```{r crossbreak}

tbl <- df %>% tables(variables = "skipmeal", crossbreak = "RSex")
tbl

```

<br> 

## Set up huxtable styling

Uses NatCen blue colour palette. 

```{r colour palette}

blue_pal <- c("#7082d4", "#8f99de", "#abb5e5", "#c7cced", "#e3e5f5")

```

## gtsummary table components

```{r}
tbl %>% names
```

<br>

```{r}

# Identify number of footnotes rows as:

# - Total number of rows in huxtable, 
# - Minus number of header rows and number of rows in table body.

tbl[["table_body"]]

```

<br>

```{r}
# Can't see where the footnote content is found
tbl[["table_styling"]] %>% names
tbl[["table_styling"]][["footnote_body"]]

```

<br>

```{r}
tbl[["call_list"]]
```

<br>

```{r}
tbl[["tbls"]]
```

<br> 

For whatever reason the code to format the caption does not work. It's annoying but not the end of the world, when adding to Excel I do it manually. 

```{r huxtable}

# df <- tbl
# align = 1

style_hux_df <- function(df, align = 1) {
  
  whether_any_sample_sizes_below_table <- df[["table_body"]] %>% pull(label) == "Unweighted sample sizes"
  starting_row_for_base_sizes <- which(whether_any_sample_sizes_below_table) - 1

  if(any(whether_any_sample_sizes_below_table)){
    df <- df %>% modify_table_body(~.x %>% mutate(label =
        case_when(row_number() == starting_row_for_base_sizes ~ " ", TRUE ~ label))
    )
  }
  
  # Extract raw caption text (guard against no caption being set at all)
  raw_caption <- df[["table_styling"]]$caption
  
  if (length(raw_caption) == 0 || is.na(raw_caption)) {
    caption_parts <- character(0)
  } else {
    caption_divs <- str_match_all(raw_caption, "<div[^>]*>(.*?)</div>")[[1]][, 2]
    caption_parts <- caption_divs %>% str_replace_all("<[^>]*>", " ") %>% str_squish()
    caption_parts <- caption_parts[caption_parts != ""]
  }
  
  if (length(caption_parts) == 0) {
      caption_parts <- " "
  }
  
  ht <- as_hux_table(df)
  caption(ht) <- NA
  
  font(ht) <- "Arial"
  font_size(ht) <- 11
  
  header_row_num <- ht %>% header_rows() %>% sum
  number_of_body_rows <- nrow(df[["table_body"]])
  first_body_row <- header_row_num + 1
  final_body_row <- header_row_num + number_of_body_rows
  first_footer_row <- final_body_row + 1
  final_footer_row <- nrow(ht)
  starting_row_for_base_sizes_ht <- starting_row_for_base_sizes + header_row_num
  
  header_rows <- 1:header_row_num
  body_rows <- first_body_row:final_body_row
  footer_rows <- first_footer_row:final_footer_row

  ht <- map_background_color(ht, by_rows("#f5f7fa", "#ffffff")) %>%
    set_background_color(row = 1:header_row_num, value = "grey90") %>% 
    set_background_color(row = footer_rows, value = "white")

  ht <- set_bold(ht, 1:header_row_num, everywhere, TRUE)

  ht <- set_top_border(ht, row = 1:final_body_row, col = everywhere, value = 1)
  ht <- set_bottom_border(ht, row = 1:final_body_row, col = everywhere, value = 1)
  ht <- set_left_border(ht, row = 1:final_body_row, col = everywhere, value = 1)
  ht <- set_right_border(ht, row = 1:final_body_row, col = everywhere, value = 1)
  ht <- set_top_border(ht, row = c(1, first_body_row), col = everywhere, value = 2)
  ht <- set_bottom_border(ht, row = c(final_body_row), col = everywhere, value = 2)
  
  if(any(whether_any_sample_sizes_below_table)){
    ht <- set_top_padding(ht, row = starting_row_for_base_sizes + header_row_num, value = 5)
    ht <- set_bottom_padding(ht, row = starting_row_for_base_sizes + header_row_num, value = 5)
  }
  
  rows_to_pad <- which(df[["table_body"]]["row_type"] == "label") + 2
  ht <- set_top_padding(ht, row = rows_to_pad, value = 3)
  ht <- set_bottom_padding(ht, row = rows_to_pad, value = 3)
  
  align_vec <- ifelse(1:ncol(ht) <= align, "left", "right")
  for (col in seq_along(align_vec)) {
    ht <- set_align(ht, everywhere, col, align_vec[col])
  }
  
  na_string(ht) <- "" 
  
  if (length(footer_rows) > 0) {
      new_footer_blocks <- list()
      for (r in footer_rows) {
        raw_text <- ht[[r, 1]]
        if (str_detect(raw_text, "<br|<b>")) {
          footer_lines <- raw_text %>% str_split_1("(?i)<br\\s*/?>") %>% trimws()
          footer_lines <- footer_lines[footer_lines != ""]
          is_bold_line <- str_detect(footer_lines, "^<b>.*</b>$")
          footer_lines_clean <- footer_lines %>% str_replace_all("(?i)</?b>", "")
          n_lines <- length(footer_lines_clean)
          new_rows <- as_hux(matrix(footer_lines_clean, nrow = n_lines, ncol = ncol(ht)))
          for (rr in seq_len(n_lines)) new_rows <- merge_across(new_rows, rr, everywhere)
          for (rr in which(is_bold_line)) new_rows <- set_bold(new_rows, rr, everywhere, TRUE)
          new_footer_blocks[[length(new_footer_blocks) + 1]] <- new_rows
        } else {
          untouched_row <- merge_across(ht[r, ], 1, everywhere)
          new_footer_blocks[[length(new_footer_blocks) + 1]] <- untouched_row
        }
      }
      combined_footer <- do.call(rbind, new_footer_blocks)
      ht <- rbind(ht[1:(min(footer_rows) - 1), ], combined_footer)
    }
  
  if (!(length(caption_parts) == 1 && caption_parts[1] == " ")) {
      n_parts <- length(caption_parts)
      empty_cols <- matrix("", nrow = n_parts, ncol = ncol(ht))
      empty_hux <- as_hux(empty_cols)
      for (r in seq_len(n_parts)) {
          empty_hux[r, 1] <- caption_parts[r]
          empty_hux <- empty_hux %>% merge_across(r, everywhere)
      }
      empty_hux <- set_header_rows(empty_hux, seq_len(n_parts), TRUE)
      ht <- rbind(empty_hux, ht)
      ht <- set_font(ht, 1:n_parts, everywhere, "Arial")
      ht <- set_font_size(ht, 1:n_parts, everywhere, 13)
      ht <- set_bold(ht, 1, everywhere, TRUE)
      if (n_parts > 1) {
          ht <- set_italic(ht, 2:n_parts, everywhere, TRUE)
      }
      ht <- set_align(ht, 1:n_parts, everywhere, "left")
      ht <- set_bottom_border(ht, row = n_parts, col = everywhere, value = 2)
  }
  
  top_border_color(ht)    <- "grey50"
  bottom_border_color(ht) <- "grey50"
  left_border_color(ht)   <- "grey50"
  right_border_color(ht)  <- "grey50"
  
  return(ht)
}
```

```{r}

writeLines(
  c("style_hux_df <- ", # Assignment needs to be added manually
  deparse(style_hux_df)), # Extract function code
  paste0(ddir, "/style_hux_df.R")) # File name is just the function name 

```


<br>

### Trial with crossbreak

```{r huxtable formatting with crossbreaks}

huxxed <- tbl %>% style_hux_df() 
tbl %>% style_hux_df()  %>% as_flextable()

```

<br> 

### With nested tables? 

```{r huxtable formatting with nested crossbreaks}

source(paste0(ddir, "nested_tables.R"))

nested_tables

```

<br>

```{r}

ntbl <- 
  df %>% nested_tables(
  variables = "skipmeal", 
  crossbreak = "RSex", 
  nest = "HEdQual3"
)

ntbl %>% style_hux_df() 
ntbl %>% style_hux_df() %>% as_flextable()

```

<br> 

## Adding tables into Excel 

### Trial 1

This is fairly successful but the auto width function in openxlsx is just not that good.

```{r huxtable to excel function}

# Set max/min column widths-----------------------------------------------------
# Set outside of function. 
options("openxlsx.minWidth" = 10) 
options("openxlsx.maxWidth" = 15) 

addh <- function(
    ht, 
    workbook, 
    sheet
    ) {
  
# Convert huxtable to workbook (starts on row 2 and col 2)
workbook_interim <- ht |> as_Workbook(Workbook = workbook, start_row = 2, start_col = 2, sheet = sheet)

# Excel cell locations
end_row <- 1 + nrow(ht)  
end_col <- 1 + ncol(ht)

# Sheet background
showGridLines(workbook_interim, sheet, showGridLines = FALSE)
# addStyle(workbook_interim, sheet = sheet, style = createStyle(fgFill = "lightgrey"), rows = 1:1000, cols = 1:1000, gridExpand = TRUE, stack = TRUE)
# 
# # Table background (could also fill the huxtable...)
# addStyle(workbook_interim, sheet = sheet, style = createStyle(fgFill = "white"), 
#          rows = 2:end_row, 
#          cols = 2:end_col, 
#          gridExpand = TRUE, stack = TRUE)

# Required to remove row heights set by as_Workbook() - these are problematic
removeRowHeights(workbook_interim, sheet, rows = 2:end_row) 

# Wrap cells -------------------------------------------------------------------

addStyle(workbook_interim, 
         sheet, 
         style = createStyle(wrapText = T), 
         cols = 2:end_col, 
         rows = 2:end_row, 
         gridExpand = TRUE, 
         stack = T)

# Remove huxtable defined column widths and apply openxlsx standards
removeColWidths(workbook_interim, sheet, cols = 2:end_col)
setColWidths(workbook_interim, sheet, cols = 2:end_col, widths = "auto")

return(workbook_interim)

}

wb <- createWorkbook()

wb <- ntbl %>% style_hux_df %>% addh(sheet = "nested table", workbook = wb)
# Save out workbook
saveWorkbook(wb, file = paste0(ddir, "3.1. Example, not working.xlsx"), overwrite = TRUE)

```

<br>

### Trial 2

The code below fairly successfully sets a sensible column width for tables. 

It's still a little hard to navigate larger tables, but perhaps there's nothing to be done on that. 

When merging 5 separate cross-break tables it might be possible to do something with colour or borders to indicate breaks between them better. 

```{r manually set col_widths}

huxt = tbl%>% style_hux_df 
rm(huxt)

.col_width <- function(huxt) { 
  
  # Identify number of header rows in gtsummary table
  no_of_header_rows <- huxt %>% header_rows %>% sum
  header_row <- no_of_header_rows
  
  # Extract first row of huxtable
  header_row_char_length <- huxt[header_row,] %>% as_vector() %>% map_int(., str_length)
  
  # Extract outcome variable length. 
  header_row_char_length[1] <- huxt[header_row+1,1] %>% as_vector() %>% map_int(., str_length)

  # Multiple by number of pixels per character
  header_proposed_width <- header_row_char_length + 3

  # Cut off of 30 imposed, and lower bound of 10. 
  header_proposed_width_cut_off <- ifelse(header_proposed_width > 30, 30, header_proposed_width)
  header_proposed_width_cut_off <- ifelse(header_proposed_width_cut_off < 10, 10, header_proposed_width_cut_off)

  # For column 1, cut off increased to 50
  if (header_proposed_width[1] > 30) {
    header_proposed_width_cut_off[1] <- ifelse(header_proposed_width[1] > 50, 50, header_proposed_width[1])
  }
  
  # And lower bound of 20. 
  if (header_proposed_width[1] < 20) {
  header_proposed_width_cut_off[1] <- ifelse(header_proposed_width_cut_off[1] < 20, 20, header_proposed_width_cut_off[1])
  }
  
return(header_proposed_width_cut_off)

} 

tbl %>% style_hux_df() %>% .col_width()




addh <- function(
    ht, 
    workbook, 
    sheet
    ) {
  
# Convert huxtable to workbook (starts on row 2 and col 2)
workbook_interim <- ht |> as_Workbook(Workbook = workbook ,start_row = 2, start_col = 2, sheet = sheet)
# Excel cell locations
end_row <- 1 + nrow(ht)  
end_col <- 1 + ncol(ht)
# Sheet background
showGridLines(workbook_interim, sheet, showGridLines = FALSE)
# Required to remove row heights set by as_Workbook() - these are problematic
removeRowHeights(workbook_interim, sheet, rows = 2:end_row) 
# Wrap cells -------------------------------------------------------------------
addStyle(workbook_interim, 
         sheet, 
         style = createStyle(wrapText = T), 
         cols = 2:end_col, 
         rows = 2:end_row, 
         gridExpand = TRUE, 
         stack = T)

total_width <- sum(ht %>% .col_width())   # <-- this line was missing

    row_line_counts <- apply(as.matrix(ht), 1, function(row_cells) {
      max(sapply(row_cells, function(cell) {
        segments <- str_split(cell, "\n")[[1]]
        sum(ceiling(pmax(nchar(segments), 1) / total_width))
      }))
    })
    
    buffer_lines <- 0.3
    row_heights <- pmax((row_line_counts + buffer_lines) * 15, 15)
    setRowHeights(workbook_interim, sheet, rows = 2:end_row, heights = row_heights)
# Remove huxtable defined column widths and apply openxlsx standards
removeColWidths(workbook_interim, sheet, cols = 2:end_col)
col_widths <- ht %>% .col_width()
setColWidths(workbook_interim, sheet, cols = 2:end_col, widths = col_widths)
setColWidths(workbook_interim, sheet, cols = 1, widths = 4)
# Format caption... 
addStyle(workbook_interim, 
         sheet, 
         style = createStyle(wrapText = T, fontName = "Arial", fontSize = 13, textDecoration = "bold"), 
         cols = 2:end_col, 
         rows = 2, 
         gridExpand = TRUE, 
         stack = T)
setRowHeights(workbook_interim, sheet, rows = 2, heights = 30)
return(workbook_interim)
}

wb <- createWorkbook()

wb <- ntbl %>% style_hux_df %>% addh(sheet = "nested table", workbook = wb)
wb <- tbl %>% style_hux_df %>% addh(sheet = "table", workbook = wb)

# Save out workbook
saveWorkbook(wb, file = paste0(ddir, "3.2. Example, working.xlsx"), overwrite = TRUE)

```

 <br>

```{r}
writeLines(
  c("addh <- ", # Assignment needs to be added manually
  deparse(addh)), # Extract function code
  paste0(ddir, "addh.R")) # File name is just the function name 

writeLines(
  c(".col_width <- ", # Assignment needs to be added manually
  deparse(.col_width)), # Extract function code
  paste0(ddir, ".col_width.R")) # File name is just the function name 

```


